A spatial analysis of the environmental correlates of COVID-19 incidence in Spain

### 2020-04-19 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.tex

The percentaje of older adults (over 65) is the second variable related
with the population structure include like control variable. Unlike the
population density the older people in Spain is concentrate in the core
of the Iberian Peninsula (see figure \ref{plotfer1}). Our hypothesis is
that in the provinces with more older adults the social contact is lower
(Prem et al. (2017)) and therefore the a negative correlation will be
expected. This hypotesis is based in the idea that the agee-structured
social contact matrix in Spain is similar to Italy (see Singh and
Adhikari (2020) Fig 1 pag 3)

@@

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/control-map-1.pdf}
\caption{\label{fig:control-map}Spatial distribution of control
variables by province (maps omits Islas Canarias)}
\end{figure}

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/plotfer1-1.pdf}
\caption{\label{Fig-quantile} Spatial distribution of Spanish density
population}
\end{figure}

@@

@@

@@

\leavevmode\hypertarget{ref-goerlich2013population}{}%
Goerlich, F.J., Cantarino, I., 2013. A population density grid for
spain. International Journal of Geographical Information Science 27,
2247--2263.

@@

@@

\leavevmode\hypertarget{ref-prem2017projecting}{}%
Prem, K., Cook, A.R., Jit, M., 2017. Projecting social contact matrices
in 152 countries using contact surveys and demographic data. PLoS
computational biology 13, e1005697.

@@

\leavevmode\hypertarget{ref-singh2020age}{}%
Singh, R., Adhikari, R., 2020. Age-structured impact of social
distancing on the covid-19 epidemic in india. arXiv preprint
arXiv:2003.12055.

>>>>

The percentaje of older adults (over 65) is the second variable that we
consider as a control. Early evidence suggest that the case rate
mortality is higher at older ages (e.g.~The Novel Coronavirus Pneumonia
Emergency Response Epidemiology Team, 2020). It is not clear that
presence of older adults necessarily translates into higher transmission
in the population. The public health tool of choice in containing the
spread of the disease has been social distancing. In this respect, the
evidence from the field of transportation is that older adults tend to
travel less frequently, for shorter distances, and have higher rates of
immobility (e.g., Roorda et al., 2010; Morency et al., 2011; Sikder and
Pinjari, 2012). In other words, older adults are, by preference or
necessity, already in a form of social isolation. Social distancing
during the pandemic may actually reinforce that condition for older
adults, as the analysis of age-structured social contact in India,
China, and Italy of Sing and Adhikari (2020) suggests. Accordingly, our
expectation is that in the provinces with higher proportions of older
adults will tend to have similarl levels of lower social contact,
particularly since age-structured social contact matrix in Spain is
similar to Italy (Prem et al., 2017).

@@

\begin{verbatim}
## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data

## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data

## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data
\end{verbatim}

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/map-controls-1.pdf}
\caption{\label{fig:map-controls}Spatial distribution of control
variables by province (maps omits Islas Canarias)}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Wsur <-}\StringTok{ }\KeywordTok{as}\NormalTok{(spdep}\OperatorTok{::}\KeywordTok{listw2mat}\NormalTok{(listw), }\StringTok{"CsparseMatrix"}\NormalTok{)}
\NormalTok{trsur <-}\StringTok{ }\NormalTok{spatialreg}\OperatorTok{::}\KeywordTok{trW}\NormalTok{(Wsur, }\DataTypeTok{type =} \StringTok{"MC"}\NormalTok{)}
\NormalTok{impacts.sur.slm_lag11 <-}\StringTok{ }\KeywordTok{impactspsur}\NormalTok{(sur.slm_lag11, }\DataTypeTok{tr =}\NormalTok{ trsur,}\DataTypeTok{R =} \DecValTok{1000}\NormalTok{)}

\NormalTok{impacts.sur.slm_lag11[[}\DecValTok{1}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Impact measures (lag, trace):
##                                       Direct   Indirect      Total
## log(GDPpc)_1                       0.6736035  0.6395634  1.3131669
## log(Older)_1                      -0.8002302 -0.7597912 -1.5600214
## log(Density)_1                     0.1662437  0.1578427  0.3240864
## Transit_1                          0.3413888  0.3241370  0.6655259
## log(Humidity_lag11)_1             -1.5578465 -1.4791220 -3.0369685
## log(Mean_Temp_lag11)_1            -1.7365494 -1.6487943 -3.3853437
## log(Sunshine_Hours_lag11 + 0.1)_1 -0.2804735 -0.2663000 -0.5467734
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{impacts.sur.slm_lag11[[}\DecValTok{2}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Impact measures (lag, trace):
##                                        Direct    Indirect      Total
## log(Density)_2                     0.06248795  0.04048188  0.1029698
## Transit_2                          0.44006548  0.28508982  0.7251553
## log(Humidity_lag11)_2             -1.06244958 -0.68829202 -1.7507416
## log(Mean_Temp_lag11)_2            -1.47760442 -0.95724386 -2.4348483
## log(Sunshine_Hours_lag11 + 0.1)_2 -0.19072953 -0.12356126 -0.3142908
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{impacts.sur.slm_lag11[[}\DecValTok{3}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Impact measures (lag, trace):
##                                        Direct    Indirect       Total
## log(Density)_3                     0.02426235  0.01927572  0.04353807
## Transit_3                          0.43033000  0.34188455  0.77221455
## log(Humidity_lag11)_3             -0.97379005 -0.77364759 -1.74743764
## log(Mean_Temp_lag11)_3            -1.39450981 -1.10789707 -2.50240689
## log(Sunshine_Hours_lag11 + 0.1)_3 -0.09683377 -0.07693158 -0.17376535
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{myimpact <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DataTypeTok{ncol =} \DecValTok{5}\NormalTok{, }\DataTypeTok{nrow =} \DecValTok{29}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{2}\OperatorTok{:}\DecValTok{30}\NormalTok{)\{}
\NormalTok{  myimpact[i}\DecValTok{-1}\NormalTok{,] <-}\StringTok{ }\NormalTok{impacts.sur.slm_lag11[[i]]}\OperatorTok{$}\NormalTok{res}\OperatorTok{$}\NormalTok{direct}
\NormalTok{\}}
\KeywordTok{matplot}\NormalTok{(myimpact, }\DataTypeTok{type=}\StringTok{'l'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/impacts-1.pdf}

@@

\begin{itemize}
\tightlist
\item
  HACE FALTA INCLUIR UN PLOT CON LAS CORRELACIONES DE LOS RESIDUOS PARA
  DARLE RELEVANCIA A LA ESTIMACION SUR
\end{itemize}

@@

\leavevmode\hypertarget{ref-Goerlich2013population}{}%
Goerlich, F.J., Cantarino, I., 2013. A population density grid for
spain. International Journal of Geographical Information Science 27,
2247--2263.

@@

\leavevmode\hypertarget{ref-Morency2011distance}{}%
Morency, C., Páez, A., Roorda, M.J., Mercado, R.G., Farber, S., 2011.
Distance traveled in three canadian cities: Spatial analysis from the
perspective of vulnerable population segments. Journal of Transport
Geography 19, 39--50.

@@

\leavevmode\hypertarget{ref-Prem2017projecting}{}%
Prem, K., Cook, A.R., Jit, M., 2017. Projecting social contact matrices
in 152 countries using contact surveys and demographic data. PLoS
computational biology 13, e1005697.

@@

\leavevmode\hypertarget{ref-Roorda2010trip}{}%
Roorda, M.J., Paez, A., Morency, C., Mercado, R., Farber, S., 2010. Trip
generation of vulnerable populations in three canadian cities: A spatial
ordered probit approach. Transportation 37, 525--548.
doi:\href{https://doi.org/10.1007/s11116-010-9263-3}{10.1007/s11116-010-9263-3}

\leavevmode\hypertarget{ref-Sikder2012immobility}{}%
Sikder, S., Pinjari, A.R., 2012. Immobility levels and mobility
preferences of the elderly in the united states evidence from 2009
national household travel survey. Transportation Research Record
137--147. doi:\href{https://doi.org/10.3141/2318-16}{10.3141/2318-16}

\leavevmode\hypertarget{ref-Singh2020age}{}%
Singh, R., Adhikari, R., 2020. Age-structured impact of social
distancing on the covid-19 epidemic in india. arXiv preprint
arXiv:2003.12055.

\leavevmode\hypertarget{ref-Novel2020epidemiological}{}%
The Novel Coronavirus Pneumonia Emergency Response Epidemiology Team,
2020. The epidemiological characteristics of an outbreak of 2019 novel
coronavirus diseases (covid-19)---china, 2020. China CDC Weekly 2,
113--122.
